---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# datastreamr.helper

<!-- badges: start -->
<!-- badges: end -->

This package aims to support users in interfacing with the <a href="https://github.com/datastreamapp/api-docs">DataStream API</a> through its <a href="https://github.com/datastreamapp/api-docs">R package</a>. See the documentation there for instructions on requesting an API key.

##

## Installation

You can install the development version of datastreamr.helper from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("p-schaefer/datastreamr.helper")
```

## Example

You can set the API Key in the script, but a more secure solution is to save the API key as an environmental variable:

```{r setAPI}
library(datastreamr.helper)

# To set API Key for the current session, use:
if (F) datastreamr::setAPIKey('xxxxxxxxxx')

# Preferably, save the API key as an environmental variable
# then there is no need to run datastreamr::setAPIKey('xxxxxxxxxx')
if (F) usethis::edit_r_environ()
# add DATASTREAM_API_KEY="xxxxxxxxxx" to the file, save, and restart R

# Setting the API key this means users never have to worry about setting the API key again
```

Get valid values for DataStream's API: 

```{r DS_valid_vals}

DS_valid <- DS_valid_vals()

# Valid fields to filter
DS_valid$valid_filter
```

```{r DS_valid_vals1}

# valid values for lookups
DS_valid$valid_lookup$ActivityMediaName
```

The functions in datastreamr.helper accept standard R operators and formats (i.e., lists and vectors) for API queries with `qs_helper()`. The `filter` argument requires a named list, with each elements name corresponding to the field to filter by (i.e., `DS_valid$valid_filter[["metadata"]]`). The associated element can be a list or vector, but the first element of that object must be a standard R operator:

```{r DS_valid_vals2}

qs_helper(
  select=c("ActivityStartDate",
           "ActivityStartTime",
           "CharacteristicName",
           "ResultUnit",
           "ResultValue",
           "ResultDetectionCondition",
           "ResultStatusID",
           "ResultComment",
           "CreateTimestamp"),
  filter=list(
    DOI=c("==","10.25976/0gvo-9d12"),
    CharacteristicName=c("%in%",c("Specific conductance","pH")),
    ActivityStartYear=c(">",2010),
    LocationId=c("==","862774"),
    ActivityStartYear=c("==","2024")
  ),
  top=5000L
)

```

The `DS_helper()` function will use `qs_helper()` and `DS_valid_vals()` internally to validate and format inputs for the datastreamr endpoints (note that `qs_helper()` does not validate inputs, only formats them). It will also split filters into separate queries if too many are specified (maximum values specied by `.max_filt`), and iterate over values of `$top` (via `.top_list`) in instances where the resulting query is too large.

```{r DS_helper}

loc <- DS_helper(
  ep="locations",
  filter=list(
    DOI=c("==","10.25976/ori9-w562"),
    CharacteristicName=c("%in%",c("Specific conductance","pH")),
    ActivityStartYear=c(">=",2020)
  )
)

rec <- DS_helper(
  ep="records",
  select=c("MonitoringLocationID",
           "ActivityStartDate",
           "CharacteristicName",
           "ResultUnit",
           "ResultValue",
           "ResultDetectionCondition",
           "ResultStatusID",
           "ResultComment"),
  filter=list(
    DOI=c("==","10.25976/ori9-w562"),
    CharacteristicName=c("%in%",c("Specific conductance","pH")),
    ActivityStartYear=c(">=",2020),
    LocationId=c("%in%", loc$Id) # Note 'Id' used here instead of 'ID'
  )
)

rec
```

Error examples:
```{r DS_helper.error1}

try(
  DS_helper(
    ep="records",
    select=c("Incorrect_Column_Name",
             "MonitoringLocationID",
             "ActivityStartDate",
             "CharacteristicName",
             "ResultUnit",
             "ResultValue",
             "ResultDetectionCondition",
             "ResultStatusID",
             "ResultComment"),
    filter=list(
      DOI=c("==","10.25976/ori9-w562"),
      CharacteristicName=c("%in%",c("Specific conductance","pH")),
      ActivityStartYear=c(">=",2020),
      LocationId=c("%in%", loc$Id) # Note 'Id' used here instead of 'ID'
    )
  )
)
```

```{r DS_helper.error2}

try(
  DS_helper(
    ep="records",
    select=c("MonitoringLocationID",
             "ActivityStartDate",
             "CharacteristicName",
             "ResultUnit",
             "ResultValue",
             "ResultDetectionCondition",
             "ResultStatusID",
             "ResultComment"),
    filter=list(
      Incorrect_Column_Name=c("==","something_wrong"),
      DOI=c("==","10.25976/ori9-w562"),
      CharacteristicName=c("%in%",c("Specific conductance","pH")),
      ActivityStartYear=c(">=",2020),
      LocationId=c("%in%", loc$Id) # Note 'Id' used here instead of 'ID'
    )
  )
)
```

```{r DS_helper.error3}

try(
  DS_helper(
    ep="records",
    select=c("MonitoringLocationID",
             "ActivityStartDate",
             "CharacteristicName",
             "ResultUnit",
             "ResultValue",
             "ResultDetectionCondition",
             "ResultStatusID",
             "ResultComment"),
    filter=list(
      DOI=c("==","10.25976/ori9-w562"),
      CharacteristicName=c("%in%",c("Specific conductance","pH","something_wrong")),
      ActivityStartYear=c(">=",2020),
      LocationId=c("%in%", loc$Id) # Note 'Id' used here instead of 'ID'
    )
  )
)
```
